---
title: "HW 1"
author:
- Havi Khurana
- Cassie Malcom
- Merly Klaas
date: "2/10/2022"
output:
  html_document:
    toc: yes
    toc_float: yes
    code_folding: hide
    highlight: kate
  pdf_document:
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      message = FALSE,
                      error = FALSE)

#load packages
pacman::p_load(tidyverse, tidytuesdayR, dpylr, ggplot2, gghighlight, ggtext, ggrepel, rio, here, colorspace)
```

```{r read-data}
transit_cost <- read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-01-05/transit_cost.csv')
```

```{r include=FALSE}
# save(transit_cost, file = "transit_cost.csv")
```

## Question 1

Data Overview
```{r include=FALSE}
str(transit_cost)
```

```{r include=FALSE}
head(transit_cost)
```

Data Cleaning
```{r include=FALSE}
tc <- transit_cost %>%
  mutate(start_year = as.integer(start_year),
         end_year = as.integer(end_year),
         real_cost = as.numeric(real_cost))

tc
```

```{r include=FALSE}
tc2 <- tc %>% rename(id = e)

tc2
```

Adding means and standard errors for real_cost variable
```{r include=FALSE}
tc3 <- tc2 %>% 
    group_by(country) %>%
    mutate(mean_rc = mean(real_cost), n=n(), sd=sd(real_cost), se=sd/sqrt(n))

tc3
```

Filter out any country with less than 3 observations
```{r include=FALSE}
tc4 <- tc3 %>%
    filter(n >= 3)

tc4
```

Join country name data
```{r include=FALSE}
#install.packages("countrycode")
library(countrycode)
country_codes <- countrycode::codelist %>%
    select(country_name = country.name.en, country = ecb)
```

```{r include=FALSE}
tc_join <- 
    merge(tc4, country_codes, by = "country")

tc_join
```

```{r include=FALSE}
tc_final <- tc_join %>%
    filter(!is.na(country))

tc_final
```

Use the transit costs data to reproduce the following plot.
<!-- There is too much space between the y-axis tick labels and the y-axis, but I've tried many things and am still having issues. -->
```{r}
p1 <- tc_final %>%
ggplot(aes(x = mean_rc, y = reorder(country_name, mean_rc))) +
  geom_point(size = 1.8, colour = "blue") +
  ggtitle("Cost to build transit systems vary across countries") +
  geom_errorbar(aes(xmin = mean_rc - se, xmax = mean_rc + se, width = 0)) +
  theme_minimal() +
  theme(plot.title.position = "plot", # easiest way to left align title
        plot.caption = element_text(hjust = 0.5, size = 10),
        panel.grid.major.y = element_blank(),
        panel.grid.minor.x = element_blank(),
        axis.text = element_text(size = 8)) +
  labs(caption = "Data provided through #tidytuesday by the Transit Costs Project") +
    scale_x_continuous(name ="Mean Real Cost (In millions of dollars)", limits = c(0, 30000), breaks = seq(0, 30000, 10000), labels=c("0","10,000","20,000", "30,000"),
                       expand = c(0,0)) +
    scale_y_discrete(name = "Country")
   
p1
```

## Question 2

Visualize the same relation, but displaying the uncertainty using an alternative method of your choosing - Multiple error bars.
```{r}
p2 <- tc_final %>%
ggplot(aes(x = mean_rc, y = reorder(country_name, mean_rc))) +
  geom_point(size = 1.8, colour = "blue") +
  ggtitle("Cost to build transit systems vary across countries") +
  geom_errorbar(aes(xmin = mean_rc + qnorm(.025)*se,
                    xmax = mean_rc + qnorm(.975)*se,
                color = "95%")) + 
  geom_errorbar(aes(xmin = mean_rc + qnorm(.05)*se,
                    xmax = mean_rc + qnorm(.95)*se,
                color = "90%")) +
  geom_errorbar(aes(xmin = mean_rc + qnorm(.1)*se,
                    xmax = mean_rc + qnorm(.9)*se,
                color = "80%"))+
  scale_color_manual("Confidence Interval",
                     values = c("#4375D3", lighten("#4375D3", .3), lighten("#4375D3", .6))) +
  theme_minimal() +
  theme(plot.title.position = "plot", # easiest way to left align title
        plot.caption = element_text(hjust = 0.5, size = 10),
        panel.grid.major.y = element_blank(),
        panel.grid.minor.x = element_blank(),
        axis.text = element_text(size = 8)) +
  labs(caption = "Data provided through #tidytuesday by the Transit Costs Project") +
    scale_x_continuous(name ="Mean Real Cost (In millions of dollars)", breaks = seq(0, 30000, 10000), labels=c("0","10,000","20,000", "30,000"),
                       expand = c(0,0)) +
    scale_y_discrete(name = "Country")
   
p2
```

## Question 3

Compute the mean length and real_cost by city. Reproduce the following plot.

```{r}
transit_cost %>% 
    group_by(country,city) %>% 
    summarise(
        n = n(),
        length_mean = mean(length, na.rm = TRUE),
        real_cost_mean = mean(as.numeric(real_cost), na.rm = TRUE) #real_cost is char
    ) %>% 
    ggplot(aes(x = length_mean, y = real_cost_mean))+
    geom_point(aes(size = n), color = "#bf35cf")+
    scale_x_log10(limits = c(-1, 200))+
    scale_y_log10(labels = scales::dollar)+
    scale_size_binned(name = "Number of transit systems", breaks = c(5, 10, 20))+
    gghighlight(country == "IN",
                unhighlighted_params = list(color = "gray80"))+
    geom_text_repel(aes(label = city),
                  min.segment.length = 0)+
  labs(
    title = "Longer transit systems tend to cost more",
    subtitle = "<span style = 'color: #bf35cf'>**India**</span> has among the most transit systems in the world",
    x = "Average length",
    y = "Average cost",
    caption = "Note the log transformation to the axes"
     )+
    theme_minimal(base_size = 14)+
    theme(
        plot.subtitle = element_markdown(),
        legend.position="bottom",
        plot.title.position = "plot"
    )


```


## Question 4

Using basically the same data, reproduce the following plot. Note you’ll need the country_name column in your dataset.

```{r}
library(ggforce)
q4 <- transit_cost %>% 
    group_by(country,city) %>% 
    summarise(
        n = n(),
        length_mean = mean(length, na.rm = TRUE),
        real_cost_mean = mean(as.numeric(real_cost), na.rm = TRUE) #real_cost is char
    ) %>% 
    mutate(
        country_name = ifelse(country == "US", "United States", NA)
    )

q4 %>% 
    ggplot(aes(x = length_mean, y = real_cost_mean))+
    geom_point(aes(size = n), color = "gray80")+
    scale_size_binned(name = "Number of transit systems", breaks = c(5, 10, 20))+
    geom_point(data = filter(q4,country == "US"), color = "#2596be", show.legend = FALSE)+
    scale_x_log10(limits = c(-1, 200))+
    scale_y_log10(labels = scales::dollar)+
    geom_mark_ellipse(aes(group = country, 
                    label = country_name),
                    data = filter(drop_na(q4),
                                  country == "US"),
                    label.colour = "#2596be",
                    con.colour = "pink",
                    color = "pink",
                    expand = unit(1, "mm"))+
    geom_label_repel(data = filter(drop_na(q4),country == "US"), 
                     aes(label = city),
                    min.segment.length = 0, 
                    color = "gray60")+
  labs(
    title = "Longer transit systems tend to cost more",
    x = "Average length",
    y = "Average cost",
    caption = "Note the log transformation to the axes"
     )+
    theme_minimal(base_size = 14)+
    theme(
        plot.subtitle = element_markdown(),
        legend.position="bottom",
        plot.title.position = "plot"
    )
```

## Question 5

Use the crime dataset to run the following code and fit the corresponding model. Note, it may take a moment to run.

```{r include=FALSE}
crime <- import(here("data", "crime.csv")) %>% 
    janitor::clean_names()
```


```{r include=FALSE}
model_data <- crime %>% 
  mutate(neighborhood_id = relevel(factor(neighborhood_id), ref = "barnum"))

m <- glm(is_crime ~ neighborhood_id, 
         data = model_data,
         family = "binomial")

# Extract the output using broom::tidy
tidied <- broom::tidy(m)

wbarnum <- tidied %>% 
  filter(term == "neighborhood_idbarnum-west")

qnorm(ppoints(20), 
      mean = wbarnum$estimate,
      sd = wbarnum$std.error)

```


```{r include=FALSE}
dizcretized <- data.frame(
    x = qnorm(ppoints(20), 
      mean = wbarnum$estimate,
      sd = wbarnum$std.error)) %>% 
        mutate( wbarnum = ifelse( x <= 0,"#8A7D98","#2CB48E" ))
```


```{r fig.height=8, fig.width=7}
#Discretized plot
ggplot(dizcretized, aes(x)) +
  geom_dotplot(aes(fill = wbarnum), binwidth = 0.039) +
     geom_vline(xintercept = 0, 
             color = "#D04344", 
             linetype = "solid", 
             size = 2) + 
    scale_fill_identity(guide = "none") +
     scale_y_continuous(name = "", 
                     breaks = NULL) +
    labs(title = "Probability of different crime rates between neighborhoods",
         subtitle = "<span style = 'color: #8A7D98'>**West Barnum**</span> compared to <span style = 'color: #2CB48E' >**Barnum**</span>",
         x = "Difference in log odds of a crime being committed",
          caption = "Each ball represents 5% probability") +
    theme_minimal() +
     theme(plot.subtitle = element_markdown(),
           text = element_text(size=15))

#ggsave(here("plots", "Discretized-plot.pdf"))
```



