---
title: "HW 1"
author:
- Havi Khurana
- Cassie Malcom
- Merly Klaas
date: "2/10/2022"
output:
  pdf_document:
    toc: yes
  html_document:
    toc: yes
    toc_float: yes
    code_folding: hide
    highlight: kate
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      message = FALSE,
                      error = FALSE)

#load packages
pacman::p_load(tidyverse, gghighlight, ggtext, ggrepel, rio, here)
```

```{r read-data}
transit_cost <- read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-01-05/transit_cost.csv')
```

## Question 1

Use the transit costs data to reproduce the following plot. 

## Question 2

Visualize the same relation, but displaying the uncertainty using an alternative method of your choosing.

## Question 3

Compute the mean length and real_cost by city. Reproduce the following plot.

```{r}
transit_cost %>% 
    group_by(country,city) %>% 
    summarise(
        n = n(),
        length_mean = mean(length, na.rm = TRUE),
        real_cost_mean = mean(as.numeric(real_cost), na.rm = TRUE) #real_cost is char
    ) %>% 
    ggplot(aes(x = length_mean, y = real_cost_mean))+
    geom_point(aes(size = n), color = "#bf35cf")+
    scale_x_log10(limits = c(-1, 200))+
    scale_y_log10(labels = scales::dollar)+
    scale_size_binned(name = "Number of transit systems", breaks = c(5, 10, 20))+
    gghighlight(country == "IN",
                unhighlighted_params = list(color = "gray80"))+
    geom_text_repel(aes(label = city),
                  min.segment.length = 0)+
  labs(
    title = "Longer transit systems tend to cost more",
    subtitle = "<span style = 'color: #bf35cf'>**India**</span> has among the most transit systems in the world",
    x = "Average length",
    y = "Average cost",
    caption = "Note the log transformation to the axes"
     )+
    theme_minimal()+
    theme(
        plot.subtitle = element_markdown(),
        legend.position="bottom",
        plot.title.position = "plot"
    )


```


## Question 4

Using basically the same data, reproduce the following plot. Note you’ll need the country_name column in your dataset.

## Question 5

Use the crime dataset to run the following code and fit the corresponding model. Note, it may take a moment to run.

```{r }
crime <- import(here("data", "crime.csv")) %>% 
    janitor::clean_names()
```


```{r}
model_data <- crime %>% 
  mutate(neighborhood_id = relevel(factor(neighborhood_id), ref = "barnum"))

m <- glm(is_crime ~ neighborhood_id, 
         data = model_data,
         family = "binomial")

# Extract the output using broom::tidy
tidied <- broom::tidy(m)

wbarnum <- tidied %>% 
  filter(term == "neighborhood_idbarnum-west")

qnorm(ppoints(20), 
      mean = wbarnum$estimate,
      sd = wbarnum$std.error)

```


```{r}
dizcretized <- data.frame(
    x = qnorm(ppoints(20), 
      mean = wbarnum$estimate,
      sd = wbarnum$std.error)) %>% 
        mutate( wbarnum = ifelse( x <= 0,"#8A7D98","#2CB48E" ))



```


```{r fig.height=8, fig.width=7}

#Discretized plot
ggplot(dizcretized, aes(x)) +
  geom_dotplot(aes(fill = wbarnum), binwidth = 0.039) +
     geom_vline(xintercept = 0, 
             color = "#D04344", 
             linetype = "solid", 
             size = 2) + 
    scale_fill_identity(guide = "none") +
     scale_y_continuous(name = "", 
                     breaks = NULL) +
    labs(title = "Probability of different crime rates between neighborhoods",
         subtitle = "<span style = 'color: #8A7D98'>**West Barnum**</span> compared to <span style = 'color: #2CB48E' >**Barnum**</span>",
         x = "Difference in log odds of a crime being committed",
          caption = "Each ball represents 5% probability") +
    theme_minimal() +
     theme(plot.subtitle = element_markdown(),
           text = element_text(size=15))
```



