---
title: "HW 1"
author: 
    - Havi Khurana
    - Cassie Malcom
    - Merly Klaas
date: "2/10/2022"
output: 
    html_document:
        toc: TRUE
        toc_float: TRUE
        code_folding: hide
        highlight: kate
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      message = FALSE,
                      error = FALSE)

#load packages
pacman::p_load(tidyverse)
```

```{r read-data}
transit_cost <- read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-01-05/transit_cost.csv')
```

## Question 1

Use the transit costs data to reproduce the following plot. 

## Question 2

Visualize the same relation, but displaying the uncertainty using an alternative method of your choosing.

## Question 3

Compute the mean length and real_cost by city. Reproduce the following plot.

## Question 4

Using basically the same data, reproduce the following plot. Note you’ll need the country_name column in your dataset.

## Question 5

Use the crime dataset to run the following code and fit the corresponding model. Note, it may take a moment to run.

